---
title: "Data Exploration For Speed Dating"
author: "DuckieNukeEm"
date: "June 16, 2016"
output: html_document
---

```{r, echo = F}
setwd("C:/Users/Cameron/OneDrive")
library(ggplot2)
data = read.csv("./Data/Kaggle Data/SpeedDating/Speed Dating Data.csv")

```

## R Markdown
GOOD MARROW PEOPLE OF THE DATING WORlD

This is my first crack at trying one of these, so this will be rough. But please do not hestiatet to provided constructive cristism for this.

Anyways, the data set, Speed Dating Data.csv, comes from Kaggle website and can be access via the following link, <https://www.kaggle.com/annavictoria/speed-dating-experiment/downloads/Speed%20Dating%20Data.csv.zip>

#Value of what people look for

Each person was asked at the start of the session what they are looking for. They were alloted 100pt and could distriubte it among 6 features. Waves 6~9 were given a 10 pt scale each, but the researchers converted into 100pt for us (THANKS GUYS :))

---
What are you Looking for in a date

```{r, echo = F}
par(mfrow=c(2,3))
hist(data[,"attr1_1"], breaks = seq(0,100,by=5),col = "blue", freq = F, border = "black",
          main = "Histogram for Attraction", xlab = "Rank", ylab = "Counts")

hist(data[,"sinc1_1"], breaks = seq(0,100,by=5),col = "blue", freq = F, border = "black",
          main = "Histogram for sincirity", xlab = "Rank", ylab = "Counts")

hist(data[,"intel1_1"], breaks = seq(0,100,by=5),col = "blue", freq = F, border = "black",
          main = "Histogram for Intelligece", xlab = "Rank", ylab = "Counts")

hist(data[,"fun1_1"], breaks = seq(0,100,by=5),col = "blue", freq = F, border = "black",
          main = "Histogram for Fun", xlab = "Rank", ylab = "Counts")

hist(data[,"amb1_1"], breaks = seq(0,100,by=5),col = "blue", freq = F, border = "black",
          main = "Histogram for Ambition", xlab = "Rank", ylab = "Counts")

hist(data[,"shar1_1"], breaks = seq(0,100,by=5),col = "blue", freq = F, border = "black",
          main = "Histogram for shared Intrest", xlab = "Rank", ylab = "Counts")
```

People are shallow and clearly rank Looks higher then intelligence, looks like Sincirity and Fun are tied for 3, and nobody gives a crap about shared intrest and ambitions. Shallow bastards!

So how do the lads vs Ladies stack up?


```{r, echo = F}
par(mfrow=c(2,3))

hist(data[data$gender == 0 ,"attr1_1"], breaks = seq(0,100,by=5),col = "pink", freq = F, border = "black",
          main = "Histogram for Attraction for Females", xlab = "Rank", ylab = "Counts")

hist(data[data$gender == 0,"sinc1_1"], breaks = seq(0,100,by=5),col = "pink", freq = F, border = "black",
          main = "Histogram for sincirity for Females", xlab = "Rank", ylab = "Counts")

hist(data[data$gender == 0,"intel1_1"], breaks = seq(0,100,by=5),col = "pink", freq = F, border = "black",
          main = "Histogram for Intelligece for Females", xlab = "Rank", ylab = "Counts")

hist(data[data$gender == 0,"fun1_1"], breaks = seq(0,100,by=5),col = "pink", freq = F, border = "black",
          main = "Histogram for Fun for Females", xlab = "Rank", ylab = "Counts")

hist(data[data$gender == 0,"amb1_1"], breaks = seq(0,100,by=5),col = "pink", freq = F, border = "black",
          main = "Histogram for Ambition for Females", xlab = "Rank", ylab = "Counts")

hist(data[data$gender == 0,"shar1_1"], breaks = seq(0,100,by=5),col = "pink", freq = F, border = "black",
          main = "Histogram for shared Intrest for Females", xlab = "Rank", ylab = "Counts")
title(main = "What are you looking for? - Just the ladies", outer = T)
```



```{r, echo = F}
par(mfrow=c(2,3))

hist(data[data$gender == 1 ,"attr1_1"], breaks = seq(0,100,by=5),col = "light blue", freq = F, border = "black",
          main = "Histogram for Attraction for Males", xlab = "Rank", ylab = "Counts")

hist(data[data$gender == 1,"sinc1_1"], breaks = seq(0,100,by=5),col = "light blue", freq = F, border = "black",
          main = "Histogram for sincirity for Males", xlab = "Rank", ylab = "Counts")

hist(data[data$gender == 1,"intel1_1"], breaks = seq(0,100,by=5),col = "light blue", freq = F, border = "black",
          main = "Histogram for Intelligece for Males", xlab = "Rank", ylab = "Counts")

hist(data[data$gender == 1,"fun1_1"], breaks = seq(0,100,by=5),col = "light blue", freq = F, border = "black",
          main = "Histogram for Fun for Males", xlab = "Rank", ylab = "Counts")

hist(data[data$gender == 1,"amb1_1"], breaks = seq(0,100,by=5),col = "light blue", freq = F, border = "black",
          main = "Histogram for Ambition for Males", xlab = "Rank", ylab = "Counts")

hist(data[data$gender == 1,"shar1_1"], breaks = seq(0,100,by=5),col = "light blue", freq = F, border = "black",
          main = "Histogram for shared Intrest for Males", xlab = "Rank", ylab = "Counts")
title(main = "What are you looking for? - only the gents", outer = T)
```

Shit I already knew, guys care more about looks than ladies do, while ladies have a slightly higher intrest in ambition and shared intrest. Ambition, that's a curious one, why would they be diffrent? OH WELL

##What do you think your date is looking for?

Intresting stuff so far, so the participants were also asked to rank what their date was looking for



```{r, echo = F}
par(mfrow=c(2,3))
hist(data[,"attr2_1"], breaks = seq(0,100,by=5),col = "blue", freq = F, border = "black",
          main = "Histogram for Attraction - Dates View", xlab = "Rank", ylab = "Counts")

hist(data[,"sinc2_1"], breaks = seq(0,100,by=5),col = "blue", freq = F, border = "black",
          main = "Histogram for sincirity - Dates View", xlab = "Rank", ylab = "Counts")

hist(data[,"intel2_1"], breaks = seq(0,100,by=5),col = "blue", freq = F, border = "black",
          main = "Histogram for Intelligece - Dates View", xlab = "Rank", ylab = "Counts")

hist(data[,"fun2_1"], breaks = seq(0,100,by=5),col = "blue", freq = F, border = "black",
          main = "Histogram for Fun - Dates View", xlab = "Rank", ylab = "Counts")

hist(data[,"amb2_1"], breaks = seq(0,100,by=5),col = "blue", freq = F, border = "black",
          main = "Histogram for Ambition - Dates View", xlab = "Rank", ylab = "Counts")

hist(data[,"shar2_1"], breaks = seq(0,100,by=5),col = "blue", freq = F, border = "black",
  main = "Histogram for shared Intrest - Dates View", xlab = "Rank", ylab = "Counts")
```


looks like the same story as previous

So how do the lads vs Ladies stack up?


```{r, echo = F}
par(mfrow=c(2,3))

hist(data[data$gender == 0 ,"attr2_1"], breaks = seq(0,100,by=5),col = "pink", freq = F, border = "black",
          main = "Histogram for Attraction for Females - Date View", xlab = "Rank", ylab = "Counts")

hist(data[data$gender == 0,"sinc2_1"], breaks = seq(0,100,by=5),col = "pink", freq = F, border = "black",
          main = "Histogram for sincirity for Females - Date View", xlab = "Rank", ylab = "Counts")

hist(data[data$gender == 0,"intel2_1"], breaks = seq(0,100,by=5),col = "pink", freq = F, border = "black",
          main = "Histogram for Intelligece for Females - Date View", xlab = "Rank", ylab = "Counts")

hist(data[data$gender == 0,"fun2_1"], breaks = seq(0,100,by=5),col = "pink", freq = F, border = "black",
          main = "Histogram for Fun for Females - Date View", xlab = "Rank", ylab = "Counts")

hist(data[data$gender == 0,"amb2_1"], breaks = seq(0,100,by=5),col = "pink", freq = F, border = "black",
          main = "Histogram for Ambition for Females - Date View", xlab = "Rank", ylab = "Counts")

hist(data[data$gender == 0,"shar2_1"], breaks = seq(0,100,by=5),col = "pink", freq = F, border = "black",
          main = "Histogram for shared Intrest for Females - Date View", xlab = "Rank", ylab = "Counts")
title(main = "What is your date looking for? - Just the ladies", outer = T)
```



```{r, echo = F}
par(mfrow=c(2,3))

hist(data[data$gender == 1 ,"attr2_1"], breaks = seq(0,100,by=5),col = "light blue", freq = F, border = "black",
          main = "Histogram for Attraction for Males  - Date View", xlab = "Rank", ylab = "Counts")

hist(data[data$gender == 1,"sinc2_1"], breaks = seq(0,100,by=5),col = "light blue", freq = F, border = "black",
          main = "Histogram for sincirity for Males  - Date View", xlab = "Rank", ylab = "Counts")

hist(data[data$gender == 1,"intel2_1"], breaks = seq(0,100,by=5),col = "light blue", freq = F, border = "black",
          main = "Histogram for Intelligece for Males  - Date View", xlab = "Rank", ylab = "Counts")

hist(data[data$gender == 1,"fun2_1"], breaks = seq(0,100,by=5),col = "light blue", freq = F, border = "black",
          main = "Histogram for Fun for Males  - Date View", xlab = "Rank", ylab = "Counts")

hist(data[data$gender == 1,"amb2_1"], breaks = seq(0,100,by=5),col = "light blue", freq = F, border = "black",
          main = "Histogram for Ambition for Males - Date View", xlab = "Rank", ylab = "Counts")

hist(data[data$gender == 1,"shar2_1"], breaks = seq(0,100,by=5),col = "light blue", freq = F, border = "black",
          main = "Histogram for shared Intrest for Males  - Date View", xlab = "Rank", ylab = "Counts")
title(main = "What is your date looking for? - only the gents", outer = T)
```


OOOO, intresting, so the gals think the guys are shallow and really only care about looks.


##How do you stack up
