---
title: "Quiz One"
output: html_document
---
### Quiz One
***
here are the specific:

To validate Module 1 and correctly answer the questions will require that you perform the following exercise first: take the code in the file named module1.R, and modify it such that:
  (1) the segmentation variable "frequency" is replaced by its log (before it is scaled)
  (2) you select a 5-segment solution instead of a 9-segment solution.

You'll be asked to answer these questions :

What is the size of the largest segment?

What is the average purchase amount of the segment which contains, on average, the customers who have made their last purchase the most recently?

Of the five following criteria to determine the best number of segments, which one is the least relevant?

Does customer #260 belongs to the same segment than customer #5920?

Looking at the average profile of segment 1, would you say that members of this segment are typically...(you'll have to choose the right proposition)

### Prep the data
***

```{r, echo = F}

data = read.delim(file = 'purchases.txt', header = FALSE, sep = '\t', dec = '.')

# Add headers and interpret the last column as a date, extract year of purchase
colnames(data) = c('customer_id', 'purchase_amount', 'date_of_purchase')
data$date_of_purchase = as.Date(data$date_of_purchase, "%Y-%m-%d")
data$days_since       = as.numeric(difftime(time1 = "2016-01-01",
                                            time2 = data$date_of_purchase,
                                            units = "days"))
library(sqldf)

# Compute recency, frequency, and average purchase amount
customers = sqldf("SELECT customer_id,
                          MIN(days_since) AS 'recency',
                          COUNT(*) AS 'frequency',
                          AVG(purchase_amount) AS 'amount'
                   FROM data GROUP BY 1")

# Copy customer data into new data frame
new_data = customers

# Remove customer id as a variable, store it as row names
row.names(new_data) = new_data$customer_id
new_data$customer_id = NULL

# Take the log-transform of the amount, and plot
new_data$amount = log(new_data$amount)

```

1) Scaling the frequency data

```{r, echo = F}
print("Before")
head(new_data$frequency)
new_data$frequency = log(new_data$frequency)
print("After")
head(new_data$frequency)

# Standardize variables
new_data = scale(new_data)
```

2) select a 5-segment solution instead of a 9-segment solution.

```{r,echo=F}
# Take a 10% sample
sample = seq(1, 18417, by = 10)
customers_sample = customers[sample, ]
new_data_sample  = new_data[sample, ]

# Compute distance metrics on standardized data
d = dist(new_data_sample)

# Perform hierarchical clustering on distance metrics
c = hclust(d, method="ward.D2")

# Cut at 9 segments
members = cutree(c, k = 5)


```

### Answer the Questions
***
What is the size of the largest segment?

```{r}
table(members)
```

What is the average purchase amount of the segment which contains, on average, the customers who have made their last purchase the most recently?

```{r}

aggregate(customers_sample[, 2:4], by = list(members), mean)
```

Of the five following criteria to determine the best number of segments, which one is the least relevant?

```{r}
```

Does customer #260 belongs to the same segment than customer #5920?
```{r}
members
```

Looking at the average profile of segment 1, would you say that members of this segment are typically...(you'll have to choose the right proposition)

```{r}
aggregate(customers_sample[, 2:4], by = list(members), mean)

```


