---
title: "Quiz 2"
output: html_document
---
### Quiz two
***
Here are the specifics
take the code in the file named module2.R, and modify it such that the managerial segment "new active" is divided into two sub-segments called "new active high" (with an average purchase amount of $100 or more) and "new active low" (with an average purchase amount of less than $100). Tips: make sure that you apply that modification to both the 2015 and 2014 segmentations, and that you modify accordingly the code that re-order the factor "segment".




### Prep the data
***
```{r}

# Load text file into local variable called 'data'
data = read.delim(file = 'purchases.txt', header = FALSE, sep = '\t', dec = '.')

# Add headers and interpret the last column as a date, extract year of purchase
colnames(data) = c('customer_id', 'purchase_amount', 'date_of_purchase')
data$date_of_purchase = as.Date(data$date_of_purchase, "%Y-%m-%d")
data$year_of_purchase = as.numeric(format(data$date_of_purchase, "%Y"))
data$days_since       = as.numeric(difftime(time1 = "2016-01-01",
                                            time2 = data$date_of_purchase,
                                            units = "days"))

# Compute key marketing indicators using SQL language
library(sqldf)

# Compute recency, frequency, and average purchase amount
customers_2015 = sqldf("SELECT customer_id,
                               MIN(days_since) AS 'recency',
                               MAX(days_since) AS 'first_purchase',
                               COUNT(*) AS 'frequency',
                               AVG(purchase_amount) AS 'amount'
                        FROM data GROUP BY 1")


# Complete segment solution using which, and exploiting previous test as input
customers_2015$segment = "NA"
customers_2015$segment[which(customers_2015$recency > 365*3)] = "inactive"
customers_2015$segment[which(customers_2015$recency <= 365*3 & customers_2015$recency > 365*2)] = "cold"
customers_2015$segment[which(customers_2015$recency <= 365*2 & customers_2015$recency > 365*1)] = "warm"
customers_2015$segment[which(customers_2015$recency <= 365)] = "active"
customers_2015$segment[which(customers_2015$segment == "warm" & customers_2015$first_purchase <= 365*2)] = "new warm"
customers_2015$segment[which(customers_2015$segment == "warm" & customers_2015$amount < 100)] = "warm low value"
customers_2015$segment[which(customers_2015$segment == "warm" & customers_2015$amount >= 100)] = "warm high value"
customers_2015$segment[which(customers_2015$segment == "active" & customers_2015$first_purchase <= 365 & customers_2015$amount < 100)] = "new active low"
customers_2015$segment[which(customers_2015$segment == "active" & customers_2015$first_purchase <= 365 & customers_2015$amount >= 100)] = "new active high"

customers_2015$segment[which(customers_2015$segment == "active" & customers_2015$amount < 100)] = "active low value"
customers_2015$segment[which(customers_2015$segment == "active" & customers_2015$amount >= 100)] = "active high value"


# Re-order factor in a way that makes sense
customers_2015$segment = factor(x = customers_2015$segment, levels = c("inactive", "cold",
                                                             "warm high value", "warm low value", "new warm",
                                                             "active high value", "active low value", "new active low", "new active high"))


customers_2014 = sqldf("SELECT customer_id,
                               MIN(days_since) - 365 AS 'recency',
                               MAX(days_since) - 365 AS 'first_purchase',
                               COUNT(*) AS 'frequency',
                               AVG(purchase_amount) AS 'amount'
                        FROM data
                        WHERE days_since > 365
                        GROUP BY 1")

# Complete segment solution using which, and exploiting previous test as input
customers_2014$segment = "NA"
customers_2014$segment[which(customers_2014$recency > 365*3)] = "inactive"
customers_2014$segment[which(customers_2014$recency <= 365*3 & customers_2014$recency > 365*2)] = "cold"
customers_2014$segment[which(customers_2014$recency <= 365*2 & customers_2014$recency > 365*1)] = "warm"
customers_2014$segment[which(customers_2014$recency <= 365)] = "active"
customers_2014$segment[which(customers_2014$segment == "warm" & customers_2014$first_purchase <= 365*2)] = "new warm"
customers_2014$segment[which(customers_2014$segment == "warm" & customers_2014$amount < 100)] = "warm low value"
customers_2014$segment[which(customers_2014$segment == "warm" & customers_2014$amount >= 100)] = "warm high value"
customers_2014$segment[which(customers_2014$segment == "active" & customers_2014$first_purchase <= 365 & customers_2014$amount < 100)] = "new active low"
customers_2014$segment[which(customers_2014$segment == "active" & customers_2014$first_purchase <= 365 & customers_2014$amount >= 100)] = "new active high"
customers_2014$segment[which(customers_2014$segment == "active" & customers_2014$amount < 100)] = "active low value"
customers_2014$segment[which(customers_2014$segment == "active" & customers_2014$amount >= 100)] = "active high value"

# Re-order factor in a way that makes sense
customers_2014$segment = factor(x = customers_2014$segment, levels = c("inactive", "cold",
                                                                       "warm high value", "warm low value", "new warm",
                                                                       "active high value", "active low value", "new active low", "new active high"))
```


### Answer the Questions
***

How many "new active low" customers were there in 2015?
```{r}
NROW(which(customers_2015$segment == "new active low"))
```

The number of "new active high" customers has increased between 2014 and 2015. What is the rate of that increase?
```{r, echo = F}
print(paste(c("2015: ", NROW(which(customers_2015$segment == "new active high"))), collapse = ""))
print(paste(c("2014: ", NROW(which(customers_2014$segment == "new active high"))), collapse = ""))
print(paste(c("Perc Change: ", NROW(which(customers_2015$segment == "new active low"))/NROW(which(customers_2014$segment == "new active low"))*100 - 100), collapse =  ""))
```

Regarding the customers who belonged to the "new warm" segment in 2014, what was there expected revenue, all things considered, in 2015?

```{r}

revenue_2015 = sqldf("SELECT customer_id, SUM(purchase_amount) AS 'revenue_2015'
                      FROM data
                      WHERE year_of_purchase = 2015
                      GROUP BY 1")

forward = merge(customers_2014, revenue_2015, all.x = TRUE)
forward$revenue_2015[is.na(forward$revenue_2015)] = 0

mean(forward[customers_2014$segment == "new warm",]$revenue_2015)

```

In terms of expected revenue, which segment groups the least profitable customers?
```{r}
aggregate(x = forward$revenue_2015, by = list(customers_2014$segment), mean)
```
Looking at segment description, what is the average purchase amount of a customer who belongs to the "new active high" segment?

```{r}

aggregate(x = customers_2014[, 2:5], by = list(customers_2014$segment), mean)

aggregate(x = customers_2015[, 2:5], by = list(customers_2015$segment), mean)

```
