---
title: "Quiz 4"
output: html_document
---
### Quiz two
***

To validate Module 4 and correctly answer the questions will require that you perform the following exercise first: take the code in the file named module4.R, and modify it such as, in the simulations, from 2016 to 2025, 1,000 new customers are acquired every year and added to the database. Tip: it will require that you add a line of code between the lines #99 and #100 to specify that segment 8 has an influx of 1,000 new customers every year.


### prepping the data
***
```{r}
# Load text file into local variable called 'data'
data = read.delim(file = 'purchases.txt', header = FALSE, sep = '\t', dec = '.')

# Add headers and interpret the last column as a date, extract year of purchase
colnames(data) = c('customer_id', 'purchase_amount', 'date_of_purchase')
data$date_of_purchase = as.Date(data$date_of_purchase, "%Y-%m-%d")
data$year_of_purchase = as.numeric(format(data$date_of_purchase, "%Y"))
data$days_since       = as.numeric(difftime(time1 = "2016-01-01",
                                            time2 = data$date_of_purchase,
                                            units = "days"))
head(data)
# Invoke library to compute key marketing indicators using SQL language
library(sqldf)

# Segment customers in 2015
customers_2015 = sqldf("SELECT customer_id,
                               MIN(days_since) AS 'recency',
                               MAX(days_since) AS 'first_purchase',
                               COUNT(*) AS 'frequency',
                               AVG(purchase_amount) AS 'amount'
                        FROM data GROUP BY 1")
customers_2015$segment = "NA"
customers_2015$segment[which(customers_2015$recency > 365*3)] = "inactive"
customers_2015$segment[which(customers_2015$recency <= 365*3 & customers_2015$recency > 365*2)] = "cold"
customers_2015$segment[which(customers_2015$recency <= 365*2 & customers_2015$recency > 365*1)] = "warm"
customers_2015$segment[which(customers_2015$recency <= 365)] = "active"
customers_2015$segment[which(customers_2015$segment == "warm" & customers_2015$first_purchase <= 365*2)] = "new warm"
customers_2015$segment[which(customers_2015$segment == "warm" & customers_2015$amount < 100)] = "warm low value"
customers_2015$segment[which(customers_2015$segment == "warm" & customers_2015$amount >= 100)] = "warm high value"
customers_2015$segment[which(customers_2015$segment == "active" & customers_2015$first_purchase <= 365)] = "new active"
customers_2015$segment[which(customers_2015$segment == "active" & customers_2015$amount < 100)] = "active low value"
customers_2015$segment[which(customers_2015$segment == "active" & customers_2015$amount >= 100)] = "active high value"
customers_2015$segment = factor(x = customers_2015$segment, levels = c("inactive", "cold",
                                                                       "warm high value", "warm low value", "new warm",
                                                                       "active high value", "active low value", "new active"))

# Segment customers in 2014
customers_2014 = sqldf("SELECT customer_id,
                               MIN(days_since) - 365 AS 'recency',
                               MAX(days_since) - 365 AS 'first_purchase',
                               COUNT(*) AS 'frequency',
                               AVG(purchase_amount) AS 'amount'
                        FROM data
                        WHERE days_since > 365
                        GROUP BY 1")
customers_2014$segment = "NA"
customers_2014$segment[which(customers_2014$recency > 365*3)] = "inactive"
customers_2014$segment[which(customers_2014$recency <= 365*3 & customers_2014$recency > 365*2)] = "cold"
customers_2014$segment[which(customers_2014$recency <= 365*2 & customers_2014$recency > 365*1)] = "warm"
customers_2014$segment[which(customers_2014$recency <= 365)] = "active"
customers_2014$segment[which(customers_2014$segment == "warm" & customers_2014$first_purchase <= 365*2)] = "new warm"
customers_2014$segment[which(customers_2014$segment == "warm" & customers_2014$amount < 100)] = "warm low value"
customers_2014$segment[which(customers_2014$segment == "warm" & customers_2014$amount >= 100)] = "warm high value"
customers_2014$segment[which(customers_2014$segment == "active" & customers_2014$first_purchase <= 365)] = "new active"
customers_2014$segment[which(customers_2014$segment == "active" & customers_2014$amount < 100)] = "active low value"
customers_2014$segment[which(customers_2014$segment == "active" & customers_2014$amount >= 100)] = "active high value"
customers_2014$segment = factor(x = customers_2014$segment, levels = c("inactive", "cold",
                                                                       "warm high value", "warm low value", "new warm",
                                                                       "active high value", "active low value", "new active"))


# --- COMPUTE TRANSITION MATRIX ----------------------------


# Compute transition matrix
new_data = merge(x = customers_2014, y = customers_2015, by = "customer_id", all.x = TRUE)
head(new_data)
transition = table(new_data$segment.x, new_data$segment.y)
print(transition)

# Divide each row by its sum
transition = transition / rowSums(transition)
print(transition)


# --- USE TRANSITION MATRIX TO MAKE PREDICTIONS ------------


# Initialize a matrix with the number of customers in each segment today and after 10 periods
segments = matrix(nrow = 8, ncol = 11)
segments[, 1] = table(customers_2015$segment)
colnames(segments) = 2015:2025
row.names(segments) = levels(customers_2015$segment)


# Compute for each an every period
for (i in 2:11) {
       segments[, i] = segments[, i-1] %*% transition 
       segments[8,i] = 1000
}
print(segments)
```

### Answering the Questions
***
How many "inactive" customers does the model predict there will be in 2025?

```{r}
segments[1,11]
```

What would the database be worth by 2025 (cumulated revenues, discounted) if a constant influx of 1,000 new customers was added every year till then?

```{r}
yearly_revenue = c(0, 0, 0, 0, 0, 323.57, 52.31, 79.17)

# Compute revenue per segment
revenue_per_segment = yearly_revenue * segments
# Compute yearly revenue
yearly_revenue = colSums(revenue_per_segment)
discount_rate = 0.10
discount = 1 / ((1 + discount_rate) ^ ((1:11) - 1))
disc_yearly_revenue = yearly_revenue * discount
disc_cumulated_revenue = cumsum(disc_yearly_revenue)
disc_cumulated_revenue - disc_cumulated_revenue[1]
```

Looking at the transition matrix, what is the likelihood of a "new warm" customer to become "cold"?

```{r}
print(transition)
```

Looking at the transition matrix, the likelihood of transitioning from "new active" to "cold" customer is 0%. Why so?

After the modification to the code, how many customers will there be in the database by 2022?
```{r}
colSums(segments)
```



