---
title: "Ch.5 Comparing Groups: Table and visulizations"
output: html_document
---

# Ch.5 Comparing Groups: Table and visulizations"

##building the base dataset

We'll, I don't feel like building this base data set so I'm joinst going to download it

```{r}
seg.df = read.csv("http://goo.gl/qw303p")

summary(seg.df)

library(psych)
describe(seg.df)
```

Okay, so there are a few groups in this data set, they are:
`r unique(seg.df$Segment)`

so, let's take a look at the income. The reason we are doing this is to get a sense, just a sense, of how income various with the various groups. 

```{r}
by(seg.df$income, seg.df$Segment, mean)

by(seg.df$income, list(seg.df$Segment, seg.df$subscribe), mean)

#or, we can use the aggregate function
aggregate(seg.df$income, list(seg.df$Segment), mean)
```


This method is all good for basic data functions, but if we start getting into more complex data, we will need a better method for this, that is where the a formula notation will work, for example we could us the following:  

```{r}
aggregate(income ~ Segment + ownHome, dat=seg.df, mean)
```
* "aggregate(income ~ Segment + ownHome, dat=seg.df, mean)" `r  aggregate(income ~ Segment + ownHome, dat=seg.df, mean) `  
* "aggregate(income ~ Segment + ownHome + subscribe, data =seg.df, mean)" `r  aggregate(income ~ Segment + ownHome + subscribe, data =seg.df, mean) `
```{r}

```
* "xtabs(kids ~ Segment, data =seg.df)"
`r xtabs(kids ~ Segment, data =seg.df)`


This can also be expanded to the grids, with the lattics packages

`r require(lattice)`
* "histogram(~subscribe | Segment, data = seg.df)"
`r histogram(~subscribe | Segment, data = seg.df)`

The pipe above is a conidtional, so it's tell lattice to create the graph conidtional on segement

We can also jazz it up a bit

* "histogram(~subscribe | Segment, data = seg.df, type="count", layout = c(4,1), col = c("burlywood","darkolivegreen"))"
`r histogram(~subscribe | Segment, data = seg.df, type="count", layout = c(4,1), col = c("burlywood","darkolivegreen"))`

Or conditoinal it on more than one variables   

* "histogram(~subscribe | Segment + ownHome, data = seg.df)"
`r histogram(~subscribe | Segment + ownHome, data = seg.df)`


###Visualization by group of continouse Data

Okay, say we want to focus on income, but cut it by segement. A bar chart is always a good bet. The problem is that we have to preprocess the data becuase barcharts show the direct value (as opposed to hist that show the COUNT)

so:
* seg.income.agg = aggregate(income ~ Segment + ownHome, data = seg.df, mean)
`r seg.income.agg = aggregate(income ~ Segment + ownHome, data = seg.df, mean)`
* barchart(income ~ Segment, data = seg.income.agg, groups = ownHome, auto.key = T, par.settings = simpleTheme(col = terrain.colors(2)))
`r barchart(income ~ Segment, data = seg.income.agg, groups = ownHome, auto.key = T, par.settings = simpleTheme(col = terrain.colors(2)))`

However, a box and whisker plot will show you more detail!  
```{r}
boxplot(income ??? Segment, data=seg.df, yaxt="n", ylab="Income ($k)")
ax.seq <- seq(from=0, to=120000, by=20000)
axis(side=2, at=ax.seq, labels=paste(ax.seq/1000, "k", sep=""), las=1)
```

As you can see the Urban Hip has lower income, but also has a lot tighter spread, while traverers have a huge spread, and in general have a higher income than everyone else.

And...of course, lattice can make a better box plot with bwplot  
* bwplot(Segment ~ income, data = seg.df, horizontal=T, xlab = "Income")
`r bwplot(Segment ~ income, data = seg.df, horizontal=T, xlab = "Income")`

(Word of caution, bw uses formula notation, but it's backwards, becuase we want to see Income (the depndent var) on segement, but the notation above is Segement on Income)


Another cool fact is we can provide conditionals as well  
* bwplot(Segment ~ income | ownHome, data = seg.df, horizontal = T, xlab = "Income")
`r bwplot(Segment ~ income | ownHome, data = seg.df, horizontal = T, xlab = "Income")`