---
title: "Ch 15 Exercises"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
#Chapter 15 Excercises

## 1
**Suppose you throw two dice. What values can the following random variables take?**  

* **The minimum face value showing**  
1,2,3,4,5 (I'm assuming no ties)

|Dice| 1 | 2 | 3 | 4 | 5 | 6 |
| -- | - | - | - | - | - | - |
| 1  | 0 | 0 | 0 | 0 | 0 | 0 |
| 2  | 1 | 0 | 0 | 0 | 0 | 0 |
| 3  | 1 | 1 | 0 | 0 | 0 | 0 |
| 4  | 1 | 1 | 1 | 0 | 0 | 0 |
| 5  | 1 | 1 | 1 | 1 | 0 | 0 |
| 6  | 1 | 1 | 1 | 1 | 1 | 0 |  
P(1) = 5/15
P(2) = 4/15
p(3) = 3/15
p(4) = 2/15
p(5) = 1/15

$\mu = 1 * \frac{5}{15} + 2 *\frac{4}{15} + 3 * \frac{3}{15} + 4 * \frac{2}{15} + 5 * \frac{1}{15} = 2.33$
* **The absolute diffrence between the face values showering**  
0,1,2,3,4,5  

|Dice| 1 | 2 | 3 | 4 | 5 | 6 |
| -- | - | - | - | - | - | - |
| 1  | 0 | 1 | 2 | 3 | 4 | 5 |
| 2  | 1 | 0 | 1 | 2 | 3 | 4 |
| 3  | 2 | 1 | 0 | 1 | 2 | 3 |
| 4  | 3 | 2 | 1 | 0 | 1 | 2 |
| 5  | 4 | 3 | 2 | 1 | 0 | 1 |
| 6  | 5 | 4 | 3 | 2 | 1 | 0 |  

p(0) = 6/36 = 1/6
p(1) = 10/36 
p(2) = 8/36
p(3) = 6/36 = 1/6
p(4) = 4/36
p(5) = 2/36 = 1/18
$\mu = ( 0* \frac{6}{36}+1*\frac{10}{36}+2*\frac{8}{36}+3*\frac{6}{36}+4*\frac{4}{36}+5*\frac{2}{36} = 1.94$

* **The Ratio; Minimum face value/other face value**  
1/6, 1/5, 1/4, 1/3, 1/2, 1/1
2/6, 2/5, 2/4, 2/3
3/6, 3/5, 3/4, 
4/6, 4/5,
5/6

|Dice| 1 | 2 | 3 | 4 | 5 | 6 |
| -- | - | - | - | - | - | - |
| 1  | 1 |1/2|1/3|1/4|1/5|1/6|
| 2  |1/2| 1 |2/3|1/2|2/5|1/3|
| 3  |1/3|2/3| 1 |3/4|3/5|1/2|
| 4  |1/4|1/2|3/4| 1 |4/5|2/3|
| 5  |1/5|2/5|3/5|4/5| 1 |5/6|
| 6  |1/6|1/3|1/2|2/3|5/6| 1 |
p(1) = 6/36 = 1/6
P(1/6) = 2/36 = 1/18
p(1/5) = 2/36 = 1/18
p(1/4) = 2/36 = 1/18
p(1/3) = 4/36 = 1/9
p(2/5) = 2/36 = 1/18
p(1/2) = 6/36 = 1/6
p(3/5) = 2/36 = 1/18
P(2/3) = 4/36 = 1/9
p(3/4) = 2/36 = 1/18
p(4/5) = 2/36 = 1/18
p(5/6) = 2/36 = 1/18

$\mu = \frac{21}{36}$


**Assuming all outcomes in the sample space are equally likely, what are the probabilitiy mass functions for these random variables? Give these in a table formate and also do a rought sketch.**

|Dice| 1 | 2 | 3 | 4 | 5 | 6 |
| -- | - | - | - | - | - | - |
| 1  |2.7|2.7|2.7|2.7|2.7|2.7|
| 2  |2.7|2.7|2.7|2.7|2.7|2.7|
| 3  |2.7|2.7|2.7|2.7|2.7|2.7|
| 4  |2.7|2.7|2.7|2.7|2.7|2.7|
| 5  |2.7|2.7|2.7|2.7|2.7|2.7|
| 6  |2.7|2.7|2.7|2.7|2.7|2.7|

**calculate the mean of each random varibale.**  

## 3

**A game consists of first tossing an unbiased coin and then rolling a sixe sided die. Let the random variable X be the score that is obtained by adding the face value of the die and the number of heads obtaine d(0 or 1). List the possible values of X and calculate it's pmf.**  

$\Omega = \{ 1,2,3,4,5,6,7\} $
p(1) = 1/12
p(2) = 1/6
p(3) = 1/6
p(4) = 1/6
p(5) = 1/6
p(6) = 1/6
p(7) = 1/12

## 5

**A discrete random variable has a pmf **$f(x)=k(\frac{1}{2})^x$ **for x = 1,2,3; f(x) = 0 for all other values of x. Find the value of k and then the mean and variance of the random variable.**

so the probability must add to 1! so we have:  

$ K*1/2 + K*1/4 + K*1/8 = 1 $  
$ 7/8 K = 1 => K = 8/7$

Ex = 8/7 * (1/2 * 1 + 1/4 * 2 + 1/8 * 3) = 8/7 * (9/8) = 9/7

((1 - 9/7)^2 + (2 - 9/7)^2 + (3-9/7)^2) = 3.530612

## 7

**Consider the discrete probability distribution defined by**
\[ p(x)=P(X=x)=\frac{1}{x(x+1)}\ for \ x=1,2,3,\dots
\]   
* **let ** $S(n)=P(X\leq n)=\sum_{x=1}^np(x)$ **. Useing the fact that **$\frac{1}{x(x+1)}=\frac{1}{x}-\frac{1}{x+1}$**, find a formula for S(n) and thus show that p is indeed a pmf.**  

Okay, so S is basically the CPD, so we need to get a formula for the CPD then take the Derivativeto get abck to the pmf.

So we have   
\[
S(n)=P(X\leq n)=\sum_{x=1}^np(x) = \sum_{x=1}^n\frac{1}{x(x+1)} =  \sum_{x=1}^n\frac{1}{x} -\sum_{x=1}^n\frac{1}{x+1}
\]  
Which, if we integrate it we get the following

\[
S = ln(x) + ln(x+1) = ln(x*(x+1))
\]

taking the dertiave yields back our original pmf. BOOM

* **write down the formula for the mean of this distribution. What is the value of this sum?**

\[ \mu = \sum_{x=1}^n \frac{x}{x*(x+1)} = \sum_{x=1}^n\frac{1}{x+1} = ....doesn't\ exists\]


## 9

**Toss a coint 20 times and let X be the length of the longest sequence of heads. We wish to estimate the probabililty function p of X. That is, for x = 1,2,...,20 we wish to estimate **$p(x)=P(X=x)$.

**Here is a function maxheads(n.toss) that sumulates X (using n.toss = 20)**

```{r}
maxheads = function(n.toss){
  n_heads = 0
  max_heads = 0
  for (i in 1:n.toss){
    if(runif(1)<0.5){
    n_heads = n_heads + 1
    } else {
    n_heads = 0
    }
    
    if(n_heads > max_heads){
    max_heads = n_heads
    }
  
  }

  return(max_heads)
}
```

**use maxheads(20) to generate an iid sample X1,....,Xn then estimate p using**
\[
\hat{p}(x) = \frac{|X_i=x|}{N} 
\]


```{r}
num_of_expirments = 100
iid = replicate(num_of_expirments, maxheads(20)) 

for( i in 1:20){
  print(paste0(i, ':', sum(i == iid)/num_of_expirments))
}
```


**print out your estimate as a table**

## 11

**suppose the rv X has continuous pdf **$f(x)=\frac{2}{x^2}\ ,\ 1\leq x\leq 2$**. Determine the mean and variance of X and find the probabililty that X exceeds 1.5**

\[ 
\mu = \int_{1}^{2} x \frac{2}{x^2} dx = \int_{1}^{2} \frac{2}{x}dx = 2ln(2) - 2ln(1) = 2ln(2)
\]

LEt's numerically test that

```{r}
to_use_fun = function(x) {2*x / x^2}
integrate(to_use_fun, 1,2)

```
*NAILED IT*

okay, let's check the variance 

\[ 
\sigma^2 = \int_{1}^{2} (x - 2log(2)) \frac{2}{x^2} dx 
\]

which I don't want to to do by hand, sooooo

```{r}
to_use_fun = function(x){2*(x - 2*(log(2))) / x^2}
integrate(to_use_fun, 1,2)

```


now what's the probl that X excided 1.5

```{r}
to_use_fun = function(x) {2 / x^2}
integrate(to_use_fun, 1,1.5)
```

so 1 less that so about 1/3