---
title: "Grey Goose Sales"
output: word_document
---
###Serving our Customers Better
We produce Grey Goose Vodka, the top super premium vodka brand in the United States. In being a super premium brand, our customer service should reflect this consideration by providing the right amount of merchandise for each of our retail sellers.

By looking at the stores in which Grey Goose is sold, we can clearly see there are five category of stores to cater to.

```{r, echo = F}
options(warn=-1)
gg <- read.csv("c:/scripts/GreyGoose.csv")
ggsales <- data.frame(aggregate(gg$TotalSales, by = list(gg$Store), FUN = sd)$x, aggregate(gg$TotalSales, by = list(gg$Store), FUN = mean))
names(ggsales) <- c("VarofSales","Store","AverageSold")
ggslesscale <- scale(ggsales[,-2])
ggdist <- dist(ggsales)
gghclust <- hclust(ggdist, method = "ward.D")
#plot(gghclust)
ggsales$Group <- cutree(gghclust, k = 5)
plot(log(VarofSales,10)~AverageSold, data = ggsales, col = Group, xlab = "Average Sales", ylab = "Log of the Varabilty of Sales", main = "Retail Groupings")
t<-aggregate(.~Group, data=ggsales[,-2], FUN = mean)
t$Group <- c("Small and Variable","Small and Steady","Large and Steady","Large and Variable","Whales")
t[,c(3,2,1)]
```

We need to focus on groups 1,4 and most importantly, 5.  

Concurrently, there seems to be no pattern to the seasonality of our Vodka. Do note that the abnormally high horizontal steak is our Group 5 Retailers from above.  

```{r, echo = F}

#gg <- read.csv("c:/scripts/GreyGoose.csv")
gg$Date <- as.Date(gg$Date)
gg$Month <- format(gg$Date, '%m')
#plot(gg$TotalSales~gg$Date, data = gg, col = Month, xlab = "Date", ylab = "Total Sales", main = "Total Sales by Day")
```
  
Crunching the numbers, we should aim for a higher production of Grey Goose Vodka during the following months: January,June, and October
```{r, echo = F} 
#summary(lm(gg$TotalSales~gg$Month, data = gg))
options(warn=0)
```